Conference Paper Critique

### 2019-03-28 conf_paper/paper.tex

\subsubsection*{Design 1}
If frequency linearity is a desired characteristic of the oscillator, an accumulator based, FPGA clock driven design exists for this use case also. 
Here the control code $k$ is added to the value stored in the accumulator at each rising edge of the FPGA clock until overflow is reached at $2^n-1$, where $n$ is the bit width of the counter, thus valuing each control code increment at $\frac{\textrm{FPGA clock}}{2^{\textrm{width}}}$ Hz. Once again the MSB of the accumulator is used as the output signal, however, with this design, the MSB will spend an equivalent duration at both logic high and low.
Again this design is better suited to lower frequencies, where the output of the DCO is orders of magnitude lower than the clock signal driving it.

\subsubsection*{Design 2}
The third and final potential DCO design is an inverter ring/chain which makes use of the delay through inverter primitives, thus bearing significant resemblance to DCOs used on ASICs. On an FPGA, the lack of precise control over the layout of blocks, means the exact value of the delay is uncertain and may vary noticeably from one implementation to another, however, this misalignment and the asynchronous behaviour stemming from the lack of an FPGA driven clock, increase the similarity between this design and an ASIC. The period is set by the number of inverters at $2\times(\textrm{num inverters})\times(\textrm{delay})$, and is varied by adding or removing a pair of inverters for every control code increment. This design works best when the desired period is within a couple orders of magnitude of the delay through each pair of inverters, beyond this point, the number of look-up tables and routing resources required to implement the design becomes unsustainable.\\

A third potential design exists, similar to Oscillator 1 in which the control code varies the reload value of the accumulator at overflow, however, this design will not be discussed in this paper.

@@

\subsubsection*{Design B}
Better suited for use with the inverter based oscillator, is a system that mimics the behaviour of a tapped delay line, working asynchronously to the FPGA clock. As the FPGA clock is not used, and that of the ADPLL cannot be used to measure itself, an unclocked sign detection circuit is required, an example of which can be seen in Figure \ref{fig:segn_pdetdl}. This is nearly identical to sign detection methods used on an ASIC, with the arbitration circuit replaced by a second SR Latch, which alleviates the issue of metastability.\\
A tapped delay line can be emulated on an FPGA through the use of inverter pairs, due to their short and approximately known delay placed between each tap.
The resolution then is the propagation time through the pair, although this will once again depend on the layout specifics, which was approximately 600 picoseconds on an Artix-7. This is significantly lower than the 4 nanoseconds achieved using the previous design and thus more suitable for use with an inverter ring oscillator, as this resolution is better than the time resolution of the oscillator. Figure \ref{fig:num_pdetdl} illustrates the configuration of the circuit used to emulate the function of a delay line that will be used in this paper.
\begin{figure}[h]
    \centering
    \includegraphics[width=0.4\textwidth]{../new_pdet1}
    \caption{Design B Sign Detection RTL Diagram.}
    \label{fig:segn_pdetdl}
    \vspace{-0.5cm}
\end{figure}
\begin{figure}[h]
    \centering
    \includegraphics[width=0.4\textwidth]{../new_pdet2}
    \caption{Tapped Delay Line Emulation RTL Diagram.}
    \label{fig:num_pdetdl}
    \vspace{-0.5cm}
\end{figure}
\section{Comparison of Designs}
This paper will now examine the performance of a number of combinations of these blocks and the performance thereof, specifically oscillator 1 with phase detector A and oscillator 2 with detectors A \& B. For each test there were four ADPLLs on the FPGA arranged in a Cartesian Grid and as such each ADPLL will be referred to using matrix indexing corresponding to their grid location. For each configuration three different modes of operation were examined: Each ADPLL locking to the same 5 MHz external reference, each ADPLL connected to the left and above neighbours in a ``uni-directional'' network and finally a true ADPLL network where all neighbours are connected. For comparison, each configuration uses loop filter gains which had previously been determined to provide the best jitter performance. The data was captured using an Agilent MSO7054-A oscilloscope at 4 GSa/s. Each point of data is the mean of five 200 microsecond long captures which were then analysed in Matlab, Table \ref{table:b_c_perf} displays the results of this analysis.\\
The first columns of the table relate to the cycle-to-cycle jitter of the system, the difference between sequential rising edges, and the second portion gives the Time Interval Error, the difference between the actual timing of the edge and the ``ideal'' time as set by the reference. This combination of measurements will show how well each design performs the task of staying consistent from cycle to cycle both in relation to itself and the system at large.

\begin{table}[!ht]
    \begin{center}
        \begin{tiny}
        	\setlength{\tabcolsep}{.9\tabcolsep}
            \begin{tabular}{cc|c|c|c|c|c|c|}           
                \cline{3-8}
                && \multicolumn{3}{c|}{Jitter Standard Deviation (ns)} & \multicolumn{3}{c|}{Max. Time Interval Error. (ns)} \T\\
                \cline{3-8} 
                &&PLL 11&PLL 12&PLL 22&PLL 11&PLL 12&PLL 22\T\\
                \hline
                \multicolumn{2}{|l|}{Osc 1, PDET A}&-&-&-&-&-&-\T\\
                \multicolumn{2}{|r|}{Free PLLs} &1.4377 &1.4714&1.4217  &6.1910&6.3667&6.8052 \T\\
                \multicolumn{2}{|r|}{Uni-dir.}  &0.91589&1.0993&1.0875  &8.3619&8.9768&10.974 \T\\
                \multicolumn{2}{|r|}{Bi-dir.}   &1.0470 &1.0979&1.0306  &24.998&33.767&33.204 \T\\
                \hline
                \multicolumn{2}{|l|}{Osc 2, PDET A}&-&-&-&-&-&-\T\\
                \multicolumn{2}{|r|}{Free PLLs} &78.458&74.025&67.779  &146.09&143.70&142.03 \T\\
                \multicolumn{2}{|r|}{Uni-dir.}  &78.428&72.646&65.665  &146.23&298.49&200.10 \T\\
                \multicolumn{2}{|r|}{Bi-dir.}   &78.436&73.827&67.648  &235.73&497.60&311.28 \T\\
                \hline
                \multicolumn{2}{|l|}{Osc 2, PDET B}&-&-&-&-&-&-\T\\
                \multicolumn{2}{|r|}{Free PLLs} &7.8090&7.8827&6.6272   &17.336&21.151&24.302 \T\\
                \multicolumn{2}{|r|}{Uni-dir.}  &7.7866&8.1145&7.3598   &17.649&29.962&37.051 \T\\
                \multicolumn{2}{|r|}{Bi-dir.}   &7.7575&7.7729&6.6962   &30.019&40.577&45.617 \T\\
                \hline
                \B
            \end{tabular}
        \end{tiny}
        \caption{}
        \label{table:b_c_perf}
    \end{center}
\vspace{-0.5cm}
\end{table}
It is apparently that the combination of the inverter chain DCO and the FPGA clocked phase detector is an ill advised choice, with the system's jitter being an order of magnitude worse than the design using inverter delay based detector. However when the resolution of the detector is then considered, a potential cause for this issue is found. The detector in case two has a resolution of 3.875 nanoseconds, as 258 MHz is the maximum FPGA clock frequency that can be used with this design, in comparison to the 1.2 nanosecond period step of the oscillator, requiring an increase in the loop filter gains. The problems of relatively inaccurate detection and increased gains combine to produce a system that is only able to achieve poor locking performance. In case three then the detector's resolution improves to 0.6 nanoseconds and produces relatively good system performance. This combination does not perform as well as that of case one, but that is attributable to the unrealistic nature of the FPGA clocked design. Both case one and three work well, delivering a level of jitter that is both low yet allows an indication of how impactful variations in architecture or control parameters are to be seen visually on an oscilloscope with appropriate persistence.\\

%From the table it is apparent that the combination of the inverter chain DCO and the FPGA clock driven phase detector is an ill-advised choice, as the jitter present in the system is several orders of magnitude worse than either other situation. In order to obtain these measurements a four times divider was inserted into the feedback path, and hence the presence of jitter greater than the period of 200 ns. In this system, the FPGA clock is at 258 MHz, the maximum possible, thus giving a detection resolution of 3.875 nanoseconds. 
%This is 3.5 times greater than the period resolution of approximately 1.2 nano second period resolution of the oscillator, and requires the gains of the PI filter to be increased fourfold in order to respond appropriately to the error detected. The 
%Both other combinations work well however, delivering a level of jitter that is both low yet allows an indication of how impactful variations in architecture or control parameters are to be seen visually on an oscilloscope.\\

>>>>

\subsubsection*{Design 1}
If frequency linearity is a desired characteristic of the oscillator, an accumulator based, FPGA clock driven design is possible. 
Here the control code $k$ is added to the value stored in the accumulator at each rising edge of the FPGA clock until overflow is reached at $2^n-1$, where $n$ is the bit width of the counter, thus valuing each control code increment at $\frac{\textrm{FPGA clock}}{2^{\textrm{width}}}$ Hz. The accumulator's Most Significant Bit (MSB) is used as the output of this block and forms a square wave.
This design is better suited to lower frequencies, where the output of the DCO is orders of magnitude lower than the clock signal driving it, as this ensures that the incremental change due to the control code remains a small fraction of the period.

\subsubsection*{Design 2}
The second potential DCO design is an inverter ring/chain which makes use of the delay through inverter primitives, thus bearing significant resemblance to DCOs used on ASICs. On an FPGA, the lack of precise control over the layout of blocks, means the exact value of the delay is uncertain and may vary noticeably from one implementation to another, however, this misalignment and the asynchronous behaviour stemming from the lack of an FPGA driven clock, increase the similarity between this design and an ASIC. The period is set by the number of inverters at $2\times(\textrm{num inverters})\times(\textrm{delay})$, and is varied by adding or removing a pair of inverters for every control code increment. This design works best when the desired period is within a couple orders of magnitude of the delay through each pair of inverters, beyond this point, the number of look-up tables and routing resources required to implement the design becomes unsustainable.\\
A third potential design exists, similar to Oscillator 1 in which the control code varies the reload value of the accumulator at overflow, however, this design will not be discussed in this paper.

@@

\subsubsection*{Design B}
Better suited for use with the inverter based oscillator, is a system that mimics the behaviour of a tapped delay line, working asynchronously to the FPGA clock. As the FPGA clock is not used, and that of the ADPLL cannot be used to measure itself, an unclocked sign detection circuit is required, an example of which can be seen in Figure \ref{fig:segn_pdetdl}. This is nearly identical to sign detection methods used on an ASIC, with the arbitration circuit replaced by a second SR Latch, which alleviates the issue of metastability.\\
A tapped delay line can be emulated on an FPGA through the use of inverter pairs, due to their short and approximately known delay placed between each tap.
The resolution then is the propagation time through the pair, although this will once again depend on the layout specifics, which was approximately 600 picoseconds on an Artix-7. This is significantly lower than the 4 nanoseconds achieved using the previous design and thus more suitable for use with an inverter ring oscillator, as this resolution is better than the time resolution of the oscillator. Figure \ref{fig:num_pdetdl} illustrates the configuration of the circuit used to emulate the function of a delay line that will be used in this paper.
\begin{figure}[h]
    \centering
    \includegraphics[width=0.4\textwidth]{../new_pdet1}
    \caption{Design B Sign Detection RTL Diagram.}
    \label{fig:segn_pdetdl}
    \vspace{-0.5cm}
\end{figure}
\begin{figure}[h]
    \centering
    \includegraphics[width=0.4\textwidth]{../new_pdet2}
    \caption{Tapped Delay Line Emulation RTL Diagram.}
    \label{fig:num_pdetdl}
    \vspace{-0.5cm}
\end{figure}
\section{Comparison of Designs}
This paper will now examine the performance of a number of combinations of these blocks and the performance thereof, specifically oscillator 1 with phase detector A and oscillator 2 with detectors A \& B. For each test there were four ADPLLs on the FPGA arranged in a Cartesian Grid and as such each ADPLL will be referred to using matrix indexing corresponding to their grid location. For each configuration three different modes of operation were examined: Each ADPLL locked to the same 5 MHz external reference, each ADPLL connected to the left and above neighbours in a ``uni-directional'' network and finally a true ADPLL network where all neighbours are connected. For comparison, each configuration uses loop filter gains which had previously been determined to provide the best jitter performance. The data was captured using an Agilent MSO7054-A oscilloscope at 4 GSa/s. Each point of data is the mean of five 200 microsecond long captures which were then analysed in Matlab, Table \ref{table:b_c_perf} displays the results of this analysis.
The first columns of the table relate to the cycle-to-cycle jitter of the system, the difference between sequential rising edges, and the second portion gives the Time Interval Error, the difference between the actual timing of the edge and the ``ideal'' time as set by the reference. This combination of measurements will show how well each design performs the task of staying consistent from cycle to cycle both in relation to itself and the system at large.

\begin{table}[!ht]
    \begin{center}
        \begin{tiny}
        	\setlength{\tabcolsep}{.9\tabcolsep}
            \begin{tabular}{cc|c|c|c|c|c|c|}           
                \cline{3-8}
                && \multicolumn{3}{c|}{Jitter Standard Deviation (ns)} & \multicolumn{3}{c|}{Max. Time Interval Error. (ns)} \T\\
                \cline{3-8} 
                &&PLL 11&PLL 12&PLL 22&PLL 11&PLL 12&PLL 22\T\\
                \hline
                \multicolumn{2}{|l|}{Osc 1, PDET A}&-&-&-&-&-&-\T\\
                \multicolumn{2}{|r|}{Free PLLs} &1.4377 &1.4714&1.4217  &6.1910&6.3667&6.8052 \T\\
                \multicolumn{2}{|r|}{Uni-dir.}  &0.91589&1.0993&1.0875  &8.3619&8.9768&10.974 \T\\
                \multicolumn{2}{|r|}{Bi-dir.}   &1.0470 &1.0979&1.0306  &24.998&33.767&33.204 \T\\
                \hline
                \multicolumn{2}{|l|}{Osc 2, PDET A}&-&-&-&-&-&-\T\\
                \multicolumn{2}{|r|}{Free PLLs} &78.458&74.025&67.779  &146.09&143.70&142.03 \T\\
                \multicolumn{2}{|r|}{Uni-dir.}  &78.428&72.646&65.665  &146.23&298.49&200.10 \T\\
                \multicolumn{2}{|r|}{Bi-dir.}   &78.436&73.827&67.648  &235.73&497.60&311.28 \T\\
                \hline
                \multicolumn{2}{|l|}{Osc 2, PDET B}&-&-&-&-&-&-\T\\
                \multicolumn{2}{|r|}{Free PLLs} &7.8090&7.8827&6.6272   &17.336&21.151&24.302 \T\\
                \multicolumn{2}{|r|}{Uni-dir.}  &7.7866&8.1145&7.3598   &17.649&29.962&37.051 \T\\
                \multicolumn{2}{|r|}{Bi-dir.}   &7.7575&7.7729&6.6962   &30.019&40.577&45.617 \T\\
                \hline
                \B
            \end{tabular}
        \end{tiny}
        \caption{}
        \label{table:b_c_perf}
    \end{center}
\vspace{-0.5cm}
\end{table}
It is apparent that the combination of the inverter chain DCO and the FPGA clocked phase detector is an ill advised choice, with the system's jitter being an order of magnitude worse than the design using inverter delay based detector. In order to obtain measurements for this combination, due to the poor performance, a four times divider was inserted into the feedback path and the comparison made between this divided output and a reference signal at 1.25 MHz, and hence the presence of jitter greater than the period of 200 ns. In this system, the FPGA clock is at 258 MHz, the maximum possible to satisfy timing requirements, thus giving a detection resolution of 3.875 nanoseconds. This is significantly greater than the 1.2 nanosecond period step of the oscillator, requiring an increase in the loop filter gains when compared to the first combination in the table. The problems of relatively inaccurate detection and increased gains combine to produce a system that is only able to achieve poor locking performance. In case three then the detector's resolution improves to 0.6 nanoseconds and produces relatively good system performance. This combination does not perform as well as that of case one, but that is attributable to the unrealistic nature of the FPGA clocked design. Both case one and three work well, delivering a level of jitter that is both low yet allows an indication of how impactful variations in architecture or control parameters are to be seen visually on an oscilloscope with appropriate persistence.\\

%From the table it is apparent that the combination of the inverter chain DCO and the FPGA clock driven phase detector is an ill-advised choice, as the jitter present in the system is several orders of magnitude worse than either other situation. 
%This is 3.5 times greater than the period resolution of approximately 1.2 nano second period resolution of the oscillator, and requires the gains of the PI filter to be increased fourfold in order to respond appropriately to the error detected. The 
%Both other combinations work well however, delivering a level of jitter that is both low yet allows an indication of how impactful variations in architecture or control parameters are to be seen visually on an oscilloscope.\\
